---
title: "VA Analysis"
editor: visual
---

title: "Virginia Election Project"
author: "Miles O'Reilly"
execute:
  echo: true
format:
  html:
    self-contained: true
    code-tools: true
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}


library(tidyverse)
library(janitor)
library(httr)
library(jsonlite)
library(kableExtra)
library(here)
library(DT)
options(scipen = 999)
options(stringsAsFactors = FALSE)



```


# Comparing Virgnia Gov vs. Prez


```{r}

joined_vacomparison <- read.csv("joined_vacomparison.csv")
DT::datatable(joined_vacomparison)

#I will calculate the margin in terms of percentage that Youngkin had within Trump voters. Positive values correlate to a higher success than Trump while negative 
Rmargin= mutate(joined_vacomparison, YoungkinMarg= pct_youngkin-trump_pct)
DT::datatable(Rmargin)

#DATA TABLE OF TOP 5 YOUNGKIN COUNTIES DIFF
Top5 <- arrange(Rmargin, desc(YoungkinMarg))
Top5county <- head(Top5,5)
DT::datatable(Top5county)

#PLOT THEM


ggplot(Top5county, aes(x=reorder(locality, YoungkinMarg), y=YoungkinMarg)) +
  geom_bar(stat = "identity", color = "black", fill = "blue") +
  coord_flip() 

# TOP 5 YOUNGKIN SUPPORT
YoungkinSenate <- arrange(joined_vacomparison, desc(pct_youngkin))
Youngkin5 <- head(YoungkinSenate, 5)
DT::datatable(Youngkin5)

#PLOT
ggplot(Youngkin5, aes(x=reorder(locality, pct_youngkin), y=pct_youngkin)) +
  geom_bar(stat = "identity", color = "black", fill = "blue") +
  coord_flip() 

#MCAULIFFE
McCauliffe <- arrange(joined_vacomparison, desc(pct_mcauliffe))
MC5 <- head(McCauliffe, 5)
DT::datatable(MC5)
  
#finding 1 - DT of highest biden margin

PresElec= mutate(joined_vacomparison, BidenMarg = biden_pct-trump_pct)
Biden5 <- arrange(PresElec, desc(BidenMarg))
Biden5county <- head(Biden5, 5)
DT::datatable(Biden5county)

#finding 2 - chart of most trump voters
trumpnumbers <- arrange(joined_vacomparison, desc(trump))
trump5 <- head(trumpnumbers, 5)
ggplot(trump5, aes(x=reorder(locality, trump), y=trump)) +
  geom_bar(stat = "identity", color = "white", fill = "red")

#finding 3 - split within democrat camp

Dmargin=mutate(joined_vacomparison, MCmarg= pct_mcauliffe-biden_pct)
Dmargincounty <-arrange(Dmargin, desc(MCmarg))
Dmargin5 <- head(Dmargincounty, 5)

#plot
ggplot(Dmargin5, aes(x=reorder(locality, MCmarg), y=MCmarg)) +
  geom_bar(stat = "identity", color = "black", fill = "blue") +
  coord_flip() 
---
