---
title: "VA Analysis"
editor: visual
---

title: "Virginia Election Project" author: "Miles O'Reilly" execute: echo: true format: html: self-contained: true code-tools: true ---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}


library(tidyverse)
library(janitor)
library(httr)
library(jsonlite)
library(kableExtra)
library(here)
library(DT)
options(scipen = 999)
options(stringsAsFactors = FALSE)



```

# Comparing Virgnia Gov vs. Prez

```{r}

joined_vacomparison <- read.csv("joined_vacomparison.csv") 

DT::datatable(joined_vacomparison)
```

```{r}
#I will calculate the margin in terms of percentage that Youngkin had within Trump voters. Positive values correlate to a higher success than Trump while negative 

Rmargin= mutate(joined_vacomparison, YoungkinMarg= pct_youngkin-trump_pct) 
DT::datatable(Rmargin)

Top5 <- arrange(Rmargin, desc(YoungkinMarg)) 
Top5county <- head(Top5,5) 
DT::datatable(Top5county)

ggplot(Top5county, aes(x=reorder(locality, YoungkinMarg), y=YoungkinMarg)) + geom_bar(stat = "identity", color = "black", fill = "blue") + coord_flip()

```

```{r}
# TOP 5 YOUNGKIN SUPPORT

YoungkinSenate <- arrange(joined_vacomparison, desc(pct_youngkin)) 
Youngkin5 <- head(YoungkinSenate, 5) 
DT::datatable(Youngkin5)

ggplot(Youngkin5, aes(x=reorder(locality, pct_youngkin), y=pct_youngkin)) + geom_bar(stat = "identity", color = "black", fill = "blue") + coord_flip()
```

```{r}
McCauliffe <- arrange(joined_vacomparison, desc(pct_mcauliffe))
MC5 <- head(McCauliffe, 5) 
DT::datatable(MC5)
```

```{r}
#finding 1 - DT of highest biden margin

PresElec= mutate(joined_vacomparison, BidenMarg = biden_pct-trump_pct) 
Biden5 <- arrange(PresElec, desc(BidenMarg)) 
Biden5county <- head(Biden5, 5) 
DT::datatable(Biden5county)

```

```{r}
#finding 2 - chart of most trump voters 

trumpnumbers <- arrange(joined_vacomparison, desc(trump))
trump5 <- head(trumpnumbers, 5) 
ggplot(trump5, aes(x=reorder(locality, trump), y=trump)) + geom_bar(stat = "identity", color = "white", fill = "red")

```

```{r}

#finding 3 - split within democrat camp

Dmargin= mutate(joined_vacomparison, MCmarg= pct_mcauliffe-biden_pct) 
Dmargincounty <-arrange(Dmargin, desc(MCmarg)) 
Dmargin5 <- head(Dmargincounty, 5)
DT::datatable(Dmargin5)

ggplot(Dmargin5, aes(x=reorder(locality, MCmarg), y=MCmarg)) + geom_bar(stat = "identity", color = "black", fill = "blue") + 
  coord_flip() 
---
